DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-01 rebuttal.tex

\mdeff{Agreed that we should do at least one thing they asked for.
\begin{itemize}
	\item Testing on the 1/12 of the sphere is not what he asked for. On one hand, a global projection would exhibit more distortion. On the other, we never use the whole sphere.
	\item Keep in mind that there is no orientation that is consistent for the 12 base faces (i.e., where is up?).
	\item Agreed that the performance should be very similar. Both are doing local computations. And locally (3 to 5 pixels across), the sphere is isomorphic to an Euclidean plane.
	\item The only difference, in my opinion, would be due to the anisotropic filters the 2D CNN has access to.
	\item Not sure we should include this result in the main paper. Hard to justify in the story, especially as a comparison with alternative spherical CNNs is missing. Let's see the results.
\end{itemize}
}

\subsection{}

\begin{mdframed}[style=comment]
Similarly, I think it would useful to check how well the DeepSphere CNN does against the FCN when the input data is rotated randomly during training - I imagine that the results should match or probably exceed the FCN results and I think this would be a neat little test which doesn't involve needing extra data, just more augmentation.
\end{mdframed}

That comparison is not warranted.
% or justified. pointless is maybe too aggressive
There are two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture of the model, or (ii) augment the dataset such that a (more general) model learns them ``brute-force''.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
Assuming the task is truly invariant to said symmetry,\footnote{As argued in the introduction, cosmological maps are rotation invariant.} the performance of the first kind of models would be exactly the same as without augmentation, while the second kind would only catch up after seeing enough data to learn the invariance.
That is a waste of resources which is only justified if it's not possible to back the desired invariance in the architecture.
The reviewer himself makes that point in comment 2.8: ``building a NN to suit your purposes''.
We updated the manuscript to more precisely reflect this argumentation.
\TK{Seems a bit strong to me.. maybe we would write it in a softer language?}

@@

- I don't see that this is general at all, only if you are considering say, preprocessed postage stamps of galaxies. I can't think of many other cases in modern machine learning where the images contain the informative part only in the centre of an image. I do agree that the informative part is often localised though. Maybe this should be updated.
\end{mdframed}

Think of the MNIST digits, cells in biomedical imaging, faces on portraits, etc.
Most photographers prefer the subject of interest to be centered.
Then, what's important is not that the informative part is only in the center, but that a fully connected layer allows a network to focus its attention on the domain.
With average pooling, all pixels are treated equally.

\subsection{}
\begin{mdframed}[style=comment]
Than can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.

- I think you mean "This" rather than "Than". I think, however, you are making a more important point here than you realise. It is not really "intrinsic data augmentation" but rather building a NN to suit your purposes. This should be done by the entire community as, at the moment, too many people blindly build their networks without considering their data and then struggle to get the best results out. The symmetries of the data should be the first thing to influence architecture of a network, exactly as you have reasoned. You might want to add a statement about how you know what your symmetries are and thus, you know how to build a well reasoned network which is more likely to work (and that other should do the same).
\end{mdframed}

Thanks for pointing out the typo, we fixed it.
We do realize it's an important point, and agree that some people apply architectures to a different problem they have been devised for without much thoughts.
A large chunk of our paper is actually about using an appropriate NN architecture.
The point you suggest us adding is actually made in the Experiments section, where we describe the data and task, and motivate the architecture choice.
We wanted the Method section to be more generic.
% Note that the task plays a role as well (i.e., two different tasks on the same data might benefit from different architectures).
% \mdeff{that's in contradiction with his comment 2.3 where he asks us to do data augmentation with the CNN}

### 2019-03-01 rebuttal.tex

- I think you mean "This" rather than "Than". I think, however, you are making a more important point here than you realise. It is not really "intrinsic data augmentation" but rather building a NN to suit your purposes. This should be done by the entire community as, at the moment, too many people blindly build their networks without considering their data and then struggle to get the best results out. The symmetries of the data should be the first thing to influence architecture of a network, exactly as you have reasoned. You might want to add a statement about how you know what your symmetries are and thus, you know how to build a well reasoned network which is more likely to work (and that other should do the same).
\end{mdframed}

Thanks for pointing out the typo, we fixed it.
We do realize it's an important point, and agree that some people apply architectures to a different problem they have been devised for without much thoughts.
A large chunk of our paper is actually about using an appropriate NN architecture.
The point you suggest us adding is actually made in the Experiments section, where we describe the data and task, and motivate the architecture choice.
We wanted the Method section to be more generic.
% Note that the task plays a role as well (i.e., two different tasks on the same data might benefit from different architectures).
% \mdeff{that's in contradiction with his comment 2.3 where he asks us to do data augmentation with the CNN}

\subsection{}
\begin{mdframed}[style=comment]
As such, the SHT is only performed once on the input (no back and forth SHT between each layer). While this clever trick lowers the complexity of the convolution to $\mathcal{O}(N_{pix})$, the non-linearity is $\mathcal{O}(N^{3/2}_{pix})$.

- I don't understand what the difference is between the convolution and the "non-linearity". I think this sentence either needs explaining further. Is the process $\mathcal{O}(N_{pix})$ or $\mathcal{O}(N^{3/2}_{pix})$?
\end{mdframed}

In \cite{kondor2018clebsch}, the convolution is performed in the spectral (harmonic) domain with a simple Hadamard product.
Instead of recomposing the signal and applying a traditional non-linearity such as the ReLU in the pixel domain, \cite{kondor2018clebsch} proposes to apply a Clebsch–Gordan product in the spectral domain.
(As a non-linear operation preserving rotation equivariance, it is considered as the non-linearity of the NN.)
Their convolution costs $\mathcal{O}(N_{pix})$, and their non-linearity costs $\mathcal{O}(N^{3/2}_{pix})$.
Stacking those layers result in a NN of overall $\mathcal{O}(N^{3/2}_{pix})$ time complexity.
We clarified the sentence.

### 2019-03-01 scnn.tex
comment 2.8
comment 2.9

There are in general two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture, or (ii) augment the dataset such that a (more general) model learns them.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
For tasks that are rotation invariant, FCNs hence need less training data as the rotation symmetry is backed in the architecture and need not be learned.
Than can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.
Moreover, FCNs can accommodate maps with a varying number of pixels.
% agnostic to the size of the input data, i.e., it can classify maps which do not have the same number of pixels.
Such global summarization as the $AV$ layer is commonly used along graph convolutions to classify graphs of varying sizes (as it is invariant to node permutation) \citep{duvenaud2015gcn, li2015gatedgnn}.
As such, CNNs should only be used if rotation invariance is undesired.

@@

Rotation equivariance was addressed by leveraging the convolution associated to the 3D rotation group SO(3), with applications to atomization energy regression and 3D model classification, alignment and retrieval \citep{cohen2018sphericalcnn,esteves2017sphericalcnn}.
The resulting convolution is performed by (i) a spherical harmonic transform (SHT), i.e., a projection on the spherical harmonics, (ii) a  multiplication in the spectral domain, and (iii) an inverse SHT.
Note the similarity with the naive graph convolution defined in \eqnref{graph_convolution_fourier}.
Likewise, the computational cost of a convolution is dominated by the two SHTs, and a naive implementation of the SHT costs $\bO(N_{pix}^2)$ operations.
Accelerated schemes however exist for some sampling sets \citep[see][for examples]{mohlenkamp1999fast, rokhlin2006fast, reinecke2013libsharp}.
The convolutions remain nevertheless expensive, limiting the practical use of this approach.
For example with HEALPix, which was designed to have a fast SHT by being iso-latitude, the computational cost of the SHT is $\bO(N_{pix}^{3/2}) = \bO(N_{side}^3) = \bO(\ell_{max}^3)$, where $\ell_{max}$ is the largest angular frequency \citep{gorski2005healpix, reinecke2013libsharp}.\footnote{All pixels are placed on $N_{ring} = 4N_{side}-1 = \bO\left(\sqrt{N_{pix}}\right)$ rings of constant latitude.
Each ring has $\bO\left(\sqrt{N_{pix}}\right)$ pixels.
Thanks to this iso-latitude property, the SHT is computed using recurrence relations for Legendre polynomials on co-latitude and fast Fourier transforms (FFTs) on longitude.
The computational cost is thus $\sqrt{N_{pix}}$ FFTs for a total cost of $\bO\left( N_{pix} \log \sqrt{N_{pix}} \right)$, plus $\sqrt{N_{pix}}$ matrix multiplications of size $\sqrt{N_{pix}}$ for a total cost of $\bO\left(N_{pix}^{3/2}\right)$ operations.}
The Clebsh-Gordan transform can be used to perform the non-linear transformation in the spectral domain \citep{kondor2018clebsch}.
As such, the SHT is only performed once on the input (no back and forth SHT between each layer).
\nati{While this clever trick lowers the complexity of the convolution (Adamar product in the spectral domain) to $\bO(N_{pix})$, the non-linearity (Clebsch–Gordan product) is $\bO(N_{pix}^{3/2})$.}
This work is another indication that there might be a $\Omega(N_{pix}^{3/2})$ computational lower bound for the proper treatment of rotation equivariance with the spherical harmonics.
In comparison, DeepSphere scales as $\bO(N_{pix})$ (see \secref{efficient_convolution}), at the expense of not being exactly equivariant (see \ref{sec:comparison_spherical_harmonics}).
Being a mathematically well-defined rotation equivariant network is the main advantage of methods based on spherical harmonics.
See \citep{kondor2018equivariance} for a rigorous treatment of convolution and equivariance in neural networks.

>>>>

There are in general two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture, or (ii) augment the dataset such that a (more general) model learns them.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
For tasks that are rotation invariant, FCNs hence need less training data as the rotation symmetry is backed in the architecture and need not be learned.
This can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.
Moreover, FCNs can accommodate maps with a varying number of pixels.
% agnostic to the size of the input data, i.e., it can classify maps which do not have the same number of pixels.
Such global summarization as the $AV$ layer is commonly used along graph convolutions to classify graphs of varying sizes (as it is invariant to node permutation) \citep{duvenaud2015gcn, li2015gatedgnn}.
As such, CNNs should only be used if rotation invariance is undesired.

@@

Rotation equivariance was addressed by leveraging the convolution associated to the 3D rotation group SO(3), with applications to atomization energy regression and 3D model classification, alignment and retrieval \citep{cohen2018sphericalcnn,esteves2017sphericalcnn}.
The resulting convolution is performed by (i) a spherical harmonic transform (SHT), i.e., a projection on the spherical harmonics, (ii) a  multiplication in the spectral domain, and (iii) an inverse SHT.
Note the similarity with the naive graph convolution defined in \eqnref{graph_convolution_fourier}.
Likewise, the computational cost of a convolution is dominated by the two SHTs, and a naive implementation of the SHT costs $\bO(N_{pix}^2)$ operations.
Accelerated schemes however exist for some sampling sets \citep[see][for examples]{mohlenkamp1999fast, rokhlin2006fast, reinecke2013libsharp}.
The convolutions remain nevertheless expensive, limiting the practical use of this approach.
For example with HEALPix, which was designed to have a fast SHT by being iso-latitude, the computational cost of the SHT is $\bO(N_{pix}^{3/2}) = \bO(N_{side}^3) = \bO(\ell_{max}^3)$, where $\ell_{max}$ is the largest angular frequency \citep{gorski2005healpix, reinecke2013libsharp}.\footnote{All pixels are placed on $N_{ring} = 4N_{side}-1 = \bO\left(\sqrt{N_{pix}}\right)$ rings of constant latitude.
Each ring has $\bO\left(\sqrt{N_{pix}}\right)$ pixels.
Thanks to this iso-latitude property, the SHT is computed using recurrence relations for Legendre polynomials on co-latitude and fast Fourier transforms (FFTs) on longitude.
The computational cost is thus $\sqrt{N_{pix}}$ FFTs for a total cost of $\bO\left( N_{pix} \log \sqrt{N_{pix}} \right)$, plus $\sqrt{N_{pix}}$ matrix multiplications of size $\sqrt{N_{pix}}$ for a total cost of $\bO\left(N_{pix}^{3/2}\right)$ operations.}
While the Clebsh-Gordan transform can be leveraged to avoid SHTs between layers (it does a non-linear transformation in the spectral domain), the transform itself costs $\bO(N_{pix}^{3/2})$, for no reduction of overall complexity \citep{kondor2018clebsch}.
This work is another indication that there might be a $\Omega(N_{pix}^{3/2})$ computational lower bound for the proper treatment of rotation equivariance with the spherical harmonics.
In comparison, DeepSphere scales as $\bO(N_{pix})$ (see \secref{efficient_convolution}), at the expense of not being exactly equivariant (see \ref{sec:comparison_spherical_harmonics}).
Being a mathematically well-defined rotation equivariant network is the main advantage of methods based on spherical harmonics.
See \citep{kondor2018equivariance} for a rigorous treatment of convolution and equivariance in neural networks.
